Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-16 paper/ms.tex
edits to abstract

We present a scalable method for Gaussian Process regression in one dimension
with a specific emphasis on large astronomical time-series data sets.
This method, which we christen \celerite, can be applied to any Gaussian Process model where the spectral
density can be expressed as any general mixture of damped sinusoid functions.
We present a mathematical description of \celerite,
its relation to the stochastically-driven, damped simple harmonic oscillator,
and details of the implementation.  We give examples of \celerite\ applied to both
simulated and actual astronomical time-series, using it to estimate the time-series
power spectrum, while simultaneously accounting for correlated noise when inferring
model parameters.  These examples include Bayesian estimation of the strength
of stellar granulation, stellar rotation periods, asteroseismic spectra, and
transiting planet light curve parameters.  The model is flexible, fast, and
general, with potential for application to a wide range of astronomical problems,
for which we provide open-source software that implements \celerite\ in
\project{C++}, \project{Python}, and \project{Julia}.

>>>>

\dfmtodo{Make this less douchey:}
In the current era of large-scale time domain astronomy, scalable methods for
probabilistic inference are of great interest.
We present a method for Gaussian Process regression in one dimension with a
computational cost that scales linearly with the size of the dataset.
This method exploits structure in the problem when the covariance function is
expressed as a mixture of complex exponentials but it does not require evenly
spaced observations or homoscedastic noise.
This form of the covariance arises naturally when the process is a mixture of
stochastically-driven damped harmonic oscillators~--~providing a physical
motivation for and interpretation of this choice~--~but we also demonstrate
that this function can also be a sufficient effective model in many other
cases.
We present a mathematical description of the method, details of the
implementation, and compare it to existing scalable Gaussian Process methods.
We demonstrate the application of this method by applying it to simulated and
real astronomical time series datasets.
These demonstrations are examples of probabilistic inference of stellar
rotation periods, asteroseismic oscillation spectra, and transiting planet
parameters.
The method is flexible, fast, and most importantly, interpretable.
There are a wide range of potential applications within astronomical data
analysis and beyond.
We provide well tested and documented open-source implementations of this
method in \project{C++}, \project{Python}, and \project{Julia}.
